---
title: "Manipulating Data using dplyr: Quiz & Exercises"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
    df_print: paged
runtime: shiny_prerendered
description: >
  Learn to manipulate data with the package dplyr.
---

```{r setup, include=FALSE}
library(learnr)
library(dplyr)
library(ggplot2)
tutorial_options(exercise.timelimit = 10)
options(max.print=1000, dplyr.summarise.inform = FALSE)
imdb <- readRDS("imdb2021.RDS")
```



## Quiz

*Question 1*

```{r Q1, echo=FALSE}
question("What dplyr function is used to create or modify columns of a dataset?",
  answer("select()", message="select is used to retrieve columns"),
  answer("filter()", message="filter is used to retrieve rows"),
  answer("arrange()", message="arrange is used to order rows"),
  answer("summarise()", message="summarise is used to create summaries"),
  answer("mutate()",correct=TRUE),
  answer("group_by()", message="group_by is used to group rows based on some columns"), allow_retry = TRUE,random_answer_order = TRUE
)
```

*Question 2*

```{r Q2, echo=FALSE}
question("Which two dplyr functions could be responsible for having an output with fewer columns than the original dataset? (select ALL that apply)",
  answer("select()", correct = TRUE),
  answer("filter()", message="with filter it's the number of rows that gets reduced"),
  answer("arrange()", message="arrange does not affect the number of rows or columns"),
  answer("summarise()", correct = TRUE),
  answer("mutate()", message="mutate can only increase the number of columns"),
  answer("group_by()", message="group_by does not affect the number of rows or columns"), allow_retry = TRUE,random_answer_order = TRUE
)
```

*Question 3*

```{r Q3, echo=FALSE}
question("Which of these commands is correct, if I want to retrieve the entries of type movie",
  answer("filter(imdb, Type=='movie')", message="there is no column 'Type' in the dataset"),
  answer("filter(imdb, type=='movie')", correct = TRUE),
  answer("filter(imdb, Type='movie')", message="there is no column 'Type' in the dataset and the single equals sign '=' is used to assign values or objects to names"),
  answer("filter(imdb, type='movie')", message="the single equals sign '=' is used to assign values or objects to names"), allow_retry = TRUE,random_answer_order = TRUE
)
```

*Question 4*

```
group_by(imdb, year)
summarise(imdb_grouped, meanVotes = mean(numVotes))


```

```{r Q4, echo=FALSE}
question("What should I add to the above command to obtain the average number of votes by year",
  answer("`arrange(numVotes)` in between the two functions", message="look at the second line - we use the data imdb_grouped, but it is not defined yet"),
  answer("`ungroup(imdb_grouped)` after the function summarise()", message="look at the second line - we use the data imdb_grouped, but it is not defined yet"),
  answer("`imdb_grouped <-` in front of the function group_by()", correct=TRUE),
  answer("`%>%` after the function group_by()", message="it could work, but we'd also need to remove the data from the function summarise()"), allow_retry = TRUE,random_answer_order = TRUE
)
```


*Question 5*

```
imdb %>%
  mutate(age=year-birthYear) %>%
    filter(age>70 & type=="movie") %>%
      group_by(year) %>%
        summarise(n=n())

```
```{r Q5, echo=FALSE}
question("What is the output of the above command (try to find the answer without running the command)",
  answer("it shows for each year, the number of movies released by a director older than 70 years", correct=TRUE),
  answer("It shows for each director older than 70 years old, the number of years where he released at least one movie", message="the 4th line is grouping the rows by year, so the function n() in the 5th line cannot be counting a number of year"),
  answer("It shows for each year, the number of directors who are currently older than 70 years old and released a movie in that year.", message="the variable age is calculated using the column year, not the value 2020, so age cannot be the 'current' age of the directors"),
  answer("An error, as n is not part of the dataset", message="There is no error in the output. n is created by the function n(), which counts the number of rows."), allow_retry = TRUE,random_answer_order = TRUE
)
```

## Exercises


**Exercise 1a. I would like to see the data for the movie "Moneyball" - produce a filter of the data showing just this movie**

```{r ex3, exercise=TRUE}

```

<!--
```{r ex3-solution}
imdb %>% 
  filter(title=="Moneyball")
  
# or

filter(imdb,title=="Moneyball")
```
-->

**Exercise 1b. That output has too many columns! I would still like to see the data on "Moneyball", but this time only showing the columns with the title, year of release, length of movie, number of votes, average rating and director**

```{r ex3b, exercise=TRUE}

```

<!--
```{r ex3b-solution}
imdb %>% 
  filter(title=="Moneyball") %>%
    select(title,year:director)
```
-->

**Exercise 2. Find the earliest year of release for each of the different types of entry in the dataset**

```{r ex2,exercise=TRUE,error=TRUE}

```

<!--
```{r ex2-solution}

imdb %>% 
  group_by(type) %>% 
   summarise(earliest_year=min(year))

```
-->


**Exercise 3. Identify and correct the four mistakes that I made in the command below, to obtain the median duration of all the movies released after the year 2000**

```{r ex1,exercise=TRUE,error=TRUE}
imdb %>%
  filter(imdb, type="movie" & year>2000) %>%
   sumarize(medianDuration = median(length)
```

<!--
```{r ex1-solution}
imdb %>%
  filter(type=="movie" & year>2000) %>%
   summarise(medianDuration = median(length))

# 1- either imdb should be removed from filter, or the first line should be deleted
# 2- the condition of the filter function needs a double equals
# 3- the function summarise was misspelled
# 4- the parenthesis of the function summarise was not closed
```
-->


**Exercise 4. Produce a list of titles which are shared across more than one different entry in the data. Order this list so that the most common title appears at the top**

```{r ex4,exercise=TRUE}

```

<!--
```{r ex4-solution}
imdb %>%
  group_by(title) %>%
   summarise(entries=n()) %>%
    filter(entries>1) %>%
     arrange(desc(entries))
```
-->




**Exercise 5a. What are the minimum, average and maximum age of a movie director releasing a movie in the imdb dataset? (you will need to add `na.rm=T` in your summary functions to deal with the entries where the year of birth of the director is missing)**

```{r ex5,exercise=TRUE}

```

<!--
```{r ex5-solution}
imdb %>%
  mutate(age=year-birthYear) %>%
    filter(type=="movie") %>%
      summarise(minAge=min(age, na.rm=T), meanAge=mean(age, na.rm=T), maxAge=max(age, na.rm=T))

```
-->

**Exercise 5b. Which director had the longest period of time between their first released movie and their final released movie**

```{r ex5b,exercise=TRUE}

```

<!--
```{r ex5b-solution}
imdb %>%
    filter(type=="movie") %>%
     group_by(director) %>%
      summarise(first_film=min(year), last_film=max(year)) %>%
       mutate(range=last_film-first_film) %>%
        filter(range==max(range))
        
```
-->


**Exercise 6. Generate a boxplot comparing the average ratings of the movies directed by Michael Bay those directed by Christopher Nolan**

```{r ex6,exercise=TRUE}

```
<!--
```{r ex6-solution}
imdb %>%
 filter(director=="Michael Bay" | director=="Christopher Nolan") %>%
   ggplot(aes(x=director, y=averageRating))+
      geom_boxplot()

```
-->


**Exercise 7. In three parts: **

1. **Find who is the worst director of romantic comedy movies (lowest average rating). Only counting directors who made at least 5 romantic comedies, that received at least 5000 votes. **
2. **Find the worst rated movie of any genre that this director has released.**
3. **Watch this movie.**

Use the first box for 1. and the second box for 2. You don't need to try to pipe all of the way through - you can use your answer from part 1 within a new set of code for part 2

Skip 3. if you find it too hard 

*(Note/Hint: We usually run this course a couple of weeks earlier in the year when it would be a thematically relevant movie to watch. Later in the year there is even less reason to watch this!)*

```{r ex7_1,exercise=TRUE}

```
<!--
```{r ex7_1-solution}
imdb %>%
  group_by(director) %>%
    filter(romance==TRUE & comedy==TRUE & type=="movie" & numVotes>5000) %>%
      summarize(n=n(), meanRating=mean(averageRating))%>%
        filter(n>=5)%>%
          filter(meanRating==min(meanRating))
```
-->

```{r ex7_2,exercise=TRUE}

```
<!--
```{r ex7_2-solution}
imdb %>%
  filter(director=="Tyler Perry" & type=="movie") %>%
    filter(averageRating==min(averageRating))
```
-->


Solutions to these exercises will be made available when the next module of the course becomes live. Check the course site for details.


## Appendix: 'imdb' dataset


For this session, we are going to use a dataset called "imdb", which we constructed from the subsets of the Internet Movie Database made available for non-commercial purposes by the IMDb team:
<a href="https://www.imdb.com/interfaces/" target="_blank">https://www.imdb.com/interfaces/</a>


It contains the following information for all the entries in the IMDB data sets which have had more than 2000 votes rating the movie on a scale of 0-10. We've also excluded those entries which had missing data about year of release, running time and director(s) was available at the time of extraction (10/11/2021):  


```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(knitr)
data.frame(Column=c("title","type","year","length","numVotes","averageRating","director","birthYear","deathYear","genre_animation","genre_action", "genre_adventure", "genre_comedy", "genre_documentary", "genre_fantasy", "genre_romance", "genre_sci_fi", "genre_thriller"),
           Description=c("popular title of the entry",
                         "type of entry: movie, short, tvMiniSeries, tvMovie, tvSeries, tvShort, tvSpecial, video or videoGame",
                         "year of release (for series, year of release of the first episode)",
                         "duration in minutes",
                         "number of votes for the average rating of the entry",
                         "IMDb's weighted average rating for the entry",
                         "director of the entry (if multiple directors, the first one was picked)",
                         "year of birth of the director",
                         "year of death of the director",
                        "the entry is of genre animation (0/1)",
                        "the entry is of genre action (0/1)",
                        "the entry is of genre adventure (0/1)",
                        "the entry is of genre comedy (0/1)",
                        "the entry is of genre documentary (0/1)",
                        "the entry is of genre fantasy (0/1)",
                        "the entry is of genre romance (0/1)",
                        "the entry is of genre science fiction (0/1)",
                        "the entry is of genre thriller (0/1)")) %>% kable()

```


The dataset has 31,529 rows. It's too much to be displayed in full here. Below are 50 rows randomly selected from the data. Spend a bit of time familiarising yourself with the structure and columns of this dataset. You can use the arrow keys to scroll right and left.

```{r,echo=FALSE}
DT::datatable(imdb[sample(1:31529,50),])

```

For the rest of this workbook, we limited the number of lines of each output to 1000 as well. So in some cases the output may be longer than what is being displayed. We figured that you and our server wouldn't mind not having the 30,000 rows of data outputted each time you try a new command or made a mistake!

## Appendix: Useful reference links  


The official dplyr documentation: <a href="https://dplyr.tidyverse.org/" target="_blank">https://dplyr.tidyverse.org/</a> 

dplyr CheatSheet:<a href="https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf" target="_blank">https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf   </a>  

Data Manipulation Tools - Rstudio video:
<a href="https://www.youtube.com/watch?v=Zc_ufg4uW4U" target="_blank">https://www.youtube.com/watch?v=Zc_ufg4uW4U   </a> 

Some documentation on subsetting r-objects using base-R: <a href="https://bookdown.org/rdpeng/rprogdatascience/subsetting-r-objects.html" target="_blank">https://bookdown.org/rdpeng/rprogdatascience/subsetting-r-objects.html </a> 
